# Final assignment, cell morphology study

*IODS final project, R script file for Analysis.*

- Data was loaded for further analysis using normal distribution, correlation, linear regression and dimensionality reduction techniques and results were interpreted.

```{r}

# set working directory
setwd("C:/Users/Sofia/Documents/Temp/Data example/")

# Access xlsx library
library(xlsx)

# read xlsx into R
p1_006 <- read.xlsx2("p1_006.xlsx",
                     sheetIndex = 1,
                     colIndex = 1,                     
                     endCol = 10)

p3_006 <- read.xlsx2("p3_006.xlsx",
                     sheetIndex = 1,
                     colIndex = 1,                     
                     endCol = 10)

p5_006 <- read.xlsx2("p5_006.xlsx",
                     sheetIndex = 1,
                     colIndex = 1,                     
                     endCol = 10)

p7_006 <- read.xlsx2("p7_006.xlsx",
                     sheetIndex = 1,
                     colIndex = 1,                     
                     endCol = 10)

# mutate p1_006 parameters to numeric
p1_006$solun.nro <- as.numeric(as.character(p1_006$solun.nro))
p1_006$ObjectAreaCh1 <- as.numeric(as.character(p1_006$ObjectAreaCh1))
p1_006$ObjectPerimCh1 <- as.numeric(as.character(p1_006$ObjectPerimCh1))
p1_006$ObjectShapeP2ACh1 <- as.numeric(as.character(p1_006$ObjectShapeP2ACh1))
p1_006$ObjectShapeLWRCh1 <- as.numeric(as.character(p1_006$ObjectShapeLWRCh1))
p1_006$ObjectShapeBFRCh1 <- as.numeric(as.character(p1_006$ObjectShapeBFRCh1))
p1_006$ObjectLengthCh1 <- as.numeric(as.character(p1_006$ObjectLengthCh1))
p1_006$ObjectWidthCh1 <- as.numeric(as.character(p1_006$ObjectWidthCh1))
p1_006$ObjectConvexHullAreaRatioCh1 <- as.numeric(as.character(p1_006$ObjectConvexHullAreaRatioCh1))
p1_006$ObjectConvexHullPerimRatioCh1 <- as.numeric(as.character(p1_006$ObjectConvexHullPerimRatioCh1))

# mutate p3_006 parameters to numeric
p3_006$solun.nro <- as.numeric(as.character(p3_006$solun.nro))
p3_006$ObjectAreaCh1 <- as.numeric(as.character(p3_006$ObjectAreaCh1))
p3_006$ObjectPerimCh1 <- as.numeric(as.character(p3_006$ObjectPerimCh1))
p3_006$ObjectShapeP2ACh1 <- as.numeric(as.character(p3_006$ObjectShapeP2ACh1))
p3_006$ObjectShapeLWRCh1 <- as.numeric(as.character(p3_006$ObjectShapeLWRCh1))
p3_006$ObjectShapeBFRCh1 <- as.numeric(as.character(p3_006$ObjectShapeBFRCh1))
p3_006$ObjectLengthCh1 <- as.numeric(as.character(p3_006$ObjectLengthCh1))
p3_006$ObjectWidthCh1 <- as.numeric(as.character(p3_006$ObjectWidthCh1))
p3_006$ObjectConvexHullAreaRatioCh1 <- as.numeric(as.character(p3_006$ObjectConvexHullAreaRatioCh1))
p3_006$ObjectConvexHullPerimRatioCh1 <- as.numeric(as.character(p3_006$ObjectConvexHullPerimRatioCh1))

# mutate p5_006 parameters to numeric
p5_006$solun.nro <- as.numeric(as.character(p5_006$solun.nro))
p5_006$ObjectAreaCh1 <- as.numeric(as.character(p5_006$ObjectAreaCh1))
p5_006$ObjectPerimCh1 <- as.numeric(as.character(p5_006$ObjectPerimCh1))
p5_006$ObjectShapeP2ACh1 <- as.numeric(as.character(p5_006$ObjectShapeP2ACh1))
p5_006$ObjectShapeLWRCh1 <- as.numeric(as.character(p5_006$ObjectShapeLWRCh1))
p5_006$ObjectShapeBFRCh1 <- as.numeric(as.character(p5_006$ObjectShapeBFRCh1))
p5_006$ObjectLengthCh1 <- as.numeric(as.character(p5_006$ObjectLengthCh1))
p5_006$ObjectWidthCh1 <- as.numeric(as.character(p5_006$ObjectWidthCh1))
p5_006$ObjectConvexHullAreaRatioCh1 <- as.numeric(as.character(p5_006$ObjectConvexHullAreaRatioCh1))
p5_006$ObjectConvexHullPerimRatioCh1 <- as.numeric(as.character(p5_006$ObjectConvexHullPerimRatioCh1))

# mutate p7_006 parameters to numeric
p7_006$solun.nro <- as.numeric(as.character(p7_006$solun.nro))
p7_006$ObjectAreaCh1 <- as.numeric(as.character(p7_006$ObjectAreaCh1))
p7_006$ObjectPerimCh1 <- as.numeric(as.character(p7_006$ObjectPerimCh1))
p7_006$ObjectShapeP2ACh1 <- as.numeric(as.character(p7_006$ObjectShapeP2ACh1))
p7_006$ObjectShapeLWRCh1 <- as.numeric(as.character(p7_006$ObjectShapeLWRCh1))
p7_006$ObjectShapeBFRCh1 <- as.numeric(as.character(p7_006$ObjectShapeBFRCh1))
p7_006$ObjectLengthCh1 <- as.numeric(as.character(p7_006$ObjectLengthCh1))
p7_006$ObjectWidthCh1 <- as.numeric(as.character(p7_006$ObjectWidthCh1))
p7_006$ObjectConvexHullAreaRatioCh1 <- as.numeric(as.character(p7_006$ObjectConvexHullAreaRatioCh1))
p7_006$ObjectConvexHullPerimRatioCh1 <- as.numeric(as.character(p7_006$ObjectConvexHullPerimRatioCh1))

# stucture of the new dataset
#str(p1_006)
dim(p1_006)
summary(p1_006)
dim(p3_006)
summary(p3_006)
dim(p5_006)
summary(p5_006)
dim(p7_006)
summary(p7_006)

# access the tidyverse libraries tidyr, dplyr
library(tidyr); library(dplyr)

# clean columns
p1_006 <- select(p1_006, -startRow)
p3_006 <- select(p3_006, -startRow)
p5_006 <- select(p5_006, -startRow)
p7_006 <- select(p7_006, -startRow)

# clean columns
p1_006 <- select(p1_006, -solun.nro)
p3_006 <- select(p3_006, -solun.nro)
p5_006 <- select(p5_006, -solun.nro)
p7_006 <- select(p7_006, -solun.nro)

```
Normal distributions and correlations between the parameters were studied for passage 1 data (p1). 

Parameters resembling normal distrubtion mainly for ObjectShapeBFRCh1 variable, other parameters have shifted distribution. 

Strong correlation between

- ObjectAreaCh1 and ObjectPerimCh1 (0.838)
- ObjectAreaCh1 and ObjectLengthCh1 (0.698)
- ObjectAreaCh1 and ObjectWidthCh1 (0.783)
- ObjectPerimCh1 and ObjectShapeP2ACh1 (0.694)
- ObjectPerimCh1 and ObjectLengthCh1 (0.933)
- ObjectPerimCh1 and ObjectWidthCh1 (0.610)
```{r}
# access the GGally and ggplot2 libraries
library(GGally)
library(ggplot2)

# advanced plot matrix with ggpairs()
ggpairs(p1_006, mapping = aes(), lower = list(combo = wrap("facethist", bins = 20)))

# Cumulative PD and Day parameter table
PD = c(19.8, 31.3, 40.0, 45.9)
Day = c(15, 28, 42, 78)

# sd and standard deviation of cell shape ObjectAreaCh1 parameter
mean_p1_006_ObjectAreaCh1 = mean(p1_006$ObjectAreaCh1)
sd_p1_006_ObjectAreaCh1 =sd(p1_006$ObjectAreaCh1)

mean_p3_006_ObjectAreaCh1 = mean(p3_006$ObjectAreaCh1)
sd_p3_006_ObjectAreaCh1 =sd(p3_006$ObjectAreaCh1)

mean_p5_006_ObjectAreaCh1 = mean(p5_006$ObjectAreaCh1)
sd_p5_006_ObjectAreaCh1 =sd(p5_006$ObjectAreaCh1)

mean_p7_006_ObjectAreaCh1 = mean(p7_006$ObjectAreaCh1)
sd_p7_006_ObjectAreaCh1 =sd(p7_006$ObjectAreaCh1)

# sd and standard deviation of cell shape ObjectPerimCh1 parameter
mean_p1_006_ObjectPerimCh1 = mean(p1_006$ObjectPerimCh1)
sd_p1_006_ObjectPerimCh1 =sd(p1_006$ObjectPerimCh1)

mean_p3_006_ObjectPerimCh1 = mean(p3_006$ObjectPerimCh1)
sd_p3_006_ObjectPerimCh1 =sd(p3_006$ObjectPerimCh1)

mean_p5_006_ObjectPerimCh1 = mean(p5_006$ObjectPerimCh1)
sd_p5_006_ObjectPerimCh1 =sd(p5_006$ObjectPerimCh1)

mean_p7_006_ObjectPerimCh1 = mean(p7_006$ObjectPerimCh1)
sd_p7_006_ObjectPerimCh1 =sd(p7_006$ObjectPerimCh1)

# Mean and standard deviation table
mean_p1357_006_ObjectAreaCh1 = c(mean_p1_006_ObjectAreaCh1, mean_p3_006_ObjectAreaCh1, mean_p5_006_ObjectAreaCh1, mean_p7_006_ObjectAreaCh1)
sd_p1357_006_ObjectAreaCh1 = c(sd_p1_006_ObjectAreaCh1, sd_p3_006_ObjectAreaCh1, sd_p5_006_ObjectAreaCh1, sd_p7_006_ObjectAreaCh1)

# Mean and standard deviation table
mean_p1357_006_ObjectPerimCh1 = c(mean_p1_006_ObjectPerimCh1, mean_p3_006_ObjectPerimCh1, mean_p5_006_ObjectPerimCh1, mean_p7_006_ObjectPerimCh1)
sd_p1357_006_ObjectPerimCh1 = c(sd_p1_006_ObjectPerimCh1, sd_p3_006_ObjectPerimCh1, sd_p5_006_ObjectPerimCh1, sd_p7_006_ObjectPerimCh1)

#Create data.frame for data
table_p1357_ObjectAreaCh1 <- data.frame(PopulationDoubling = PD,
                  CultureTime = Day,
                  mean_ObjectAreaCh1 = mean_p1357_006_ObjectAreaCh1,
                  sd_ObjectAreaCh1 = sd_p1357_006_ObjectAreaCh1)

table_p1357_ObjectPerimCh1 <- data.frame(PopulationDoubling = PD,
                          CultureTime = Day,
                          mean_ObjectPerimCh1 = mean_p1357_006_ObjectPerimCh1,
                          sd_ObjectPerimCh1 = sd_p1357_006_ObjectPerimCh1)
```
ObjectAreaCh1 parameter was plotted against PD (Population Doubling), cell area is increasing as a function of PD.

ObjectAreaCh1 parameter was plotted against Culture Time (Days), cell area is increasing as a function of Culture Time.

ObjectPerimCh1 parameter was plotted against PD (Population Doubling), cell perimeter is increasing as a function of PD.

ObjectPerimCh1 parameter was plotted against Culture Time (Days), cell perimeter is increasing as a function of Culture Time.
```{r}
# Draw plots
ggplot(table_p1357_ObjectAreaCh1, aes(x=PopulationDoubling, y=mean_ObjectAreaCh1)) +
  geom_errorbar(aes(ymin=mean_ObjectAreaCh1-sd_ObjectAreaCh1, ymax=mean_ObjectAreaCh1+sd_ObjectAreaCh1), width=.2) +
  geom_line() +
  geom_point()

ggplot(table_p1357_ObjectAreaCh1, aes(x=CultureTime, y=mean_ObjectAreaCh1)) +
  geom_errorbar(aes(ymin=mean_ObjectAreaCh1-sd_ObjectAreaCh1, ymax=mean_ObjectAreaCh1+sd_ObjectAreaCh1), width=.2) +
  geom_line() +
  geom_point()

ggplot(table_p1357_ObjectPerimCh1, aes(x=PopulationDoubling, y=mean_ObjectPerimCh1)) +
  geom_errorbar(aes(ymin=mean_ObjectPerimCh1-sd_ObjectPerimCh1, ymax=mean_ObjectPerimCh1+sd_ObjectPerimCh1), width=.2) +
  geom_line() +
  geom_point()

ggplot(table_p1357_ObjectPerimCh1, aes(x=CultureTime, y=mean_ObjectPerimCh1)) +
  geom_errorbar(aes(ymin=mean_ObjectPerimCh1-sd_ObjectPerimCh1, ymax=mean_ObjectPerimCh1+sd_ObjectPerimCh1), width=.2) +
  geom_line() +
  geom_point()

# calculate the correlation matrix and round it
cor_matrix_p1_006 <- cor(p1_006) 
cor_matrix_p3_006 <- cor(p3_006) 
cor_matrix_p5_006 <- cor(p5_006) 
cor_matrix_p7_006 <- cor(p7_006) 

# print the correlation matrix
cor_matrix_p1_006 %>% round(digits = 2)
cor_matrix_p3_006 %>% round(digits = 2)
cor_matrix_p5_006 %>% round(digits = 2)
cor_matrix_p7_006 %>% round(digits = 2)

# access the corrplot libraries
library(corrplot)

```
Correlations between the parameters were studied for passage 1 data (p1).

Strong positive correlation between

- ObjectAreaCh1 and ObjectPerimCh1
- ObjectAreaCh1 and ObjectLengthCh1
- ObjectAreaCh1 and ObjectWidthCh1
- ObjectPerimCh1 and ObjectShapeP2ACh1
- ObjectPerimCh1 and ObjectLengthCh1
- ObjectPerimCh1 and ObjectWidthCh1

Strong negative correlation between

- ObjectAreaCh1 and ObjectShapeBFRCh1
- ObjectAreaCh1 and ObjectConvexHullPerimRatioCh1
- ObjectPerimCh1 and ObjectShapeBFRCh1
- ObjectPerimCh1 and ObjectConvexHullPerimRatioCh1

For passages 3, 5 and 7 (p3, p5 and p7) in addition of p1 correlations also more correlations found between other parameters.

```{r}
# visualize the correlation matrix
corrplot(cor_matrix_p1_006, method="circle", type = "upper", cl.pos = "b", tl.pos = "d", tl.cex = 0.6, title = "p1_006", mar=c(0,0,1,0))
corrplot(cor_matrix_p3_006, method="circle", type = "upper", cl.pos = "b", tl.pos = "d", tl.cex = 0.6, title = "p3_006", mar=c(0,0,1,0))
corrplot(cor_matrix_p5_006, method="circle", type = "upper", cl.pos = "b", tl.pos = "d", tl.cex = 0.6, title = "p5_006", mar=c(0,0,1,0))
corrplot(cor_matrix_p7_006, method="circle", type = "upper", cl.pos = "b", tl.pos = "d", tl.cex = 0.6, title = "p7_006", mar=c(0,0,1,0))

# regression model with multiple explanatory variables
my_model_p1_006 <- lm(p1_006$ObjectAreaCh1 ~ p1_006$ObjectPerimCh1 + p1_006$ObjectLengthCh1 + p1_006$ObjectWidthCh1, data = p1_006)

# print out a summary of the model
summary(my_model_p1_006)

# The relationship between the chosen explanatory variables and the target variable is statistically significant for variable p1_006$ObjectLengthCh1 and p1_006$ObjectWidthCh1 ('***' 0.001), and for p1_006$ObjectPerimCh1 ('*' 0.05). Multiple R squared of the model is 0.8307, which is high thus linear model explains well this data. (Ref Wikipedia: In statistics, the coefficient of multiple correlation is a measure of how well a given variable can be predicted using a linear function of a set of other variables).

# Diagnostic plots using the plot() function. Residuals vs Fitted values, Normal QQ-plot and Residuals vs Leverage, plots 1, 2 and 5. Some outliers seen from diagnostic plots, e.g. in Normal Q-Q plot data not linear in the beginnig and at the end points.
#par(mfrow = c(2,2))
```
Data was further analysed using linear regression model for passage 1 data (p1). The relationship between the chosen explanatory variables and the target variable is statistically significant for variable ObjectLengthCh1 and ObjectWidthCh1 (0.001), and for ObjectPerimCh1 (0.05). Multiple R squared of the model is 0.8307, which is high thus linear model explains well this data.
```{r}
plot(my_model_p1_006, which = c(1,2,5))

# principal component analysis (PCA) for standardized data

# standardize the variables
p1_006_std <- scale(p1_006)

# perform principal component analysis (with the SVD method)
pca_p1_006 <- prcomp(p1_006_std)

# draw a biplot of the principal component representation and the original variables
#biplot(pca_p1_006, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))

# create and print out a summary of pca_p1_006
s <- summary(pca_p1_006)
s

# rounded percetanges of variance captured by each PC
pca_pr <- round(100*s$importance[2, ], digits = 1)

# print out the percentages of variance
pca_pr

# create object pc_lab to be used as axis labels
pc_lab <- paste0(names(pca_pr), " (", pca_pr, "%)")
```
Data was further analysed using dimensionality reduction techniques, principal component analysis (PCA) for passage 1 data (p1). PC1 explains 47.1%  and PC2 14.6% of the variance.
```{r}
# draw a biplot
biplot(pca_p1_006, cex = c(0.8, 1), col = c("grey40", "deeppink2"), xlab = pc_lab[1], ylab = pc_lab[2])

```

